### 2015-02-03 paper1/paper1_distribution.tex

\section{\agb\, Candidate Distribution}
\label{sec:distribution}
Put words here